---
output: word_document
---
Lesson 3
========================================================

***

### What to Do First?
Notes:

```{r}
getwd()
setwd('C:/Users/Nicolas/Desktop/Projets Tech/Data Analysis with R/')
list.files()
```


***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
qplot(x= dob_day, data = pf) +
  scale_x_continuous(breaks=1:31)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
qplot(x= dob_day, data = pf) +
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month, ncol=3)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf)
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
#qplot(x = friend_count, data = pf, xlim = c(0, 1000) )
# or
qplot(x = friend_count, data = pf) +
  scale_x_continuous( limits = c(0,1000))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +    # update the x-axis label
  facet_wrap(~gender)   
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
#qplot(x = friend_count, data = na.omit(pf), binwidth = 25) +
#  scale_x_continuous(limits = c(0, 1000),
#                     breaks = seq(0, 1000, 50)) +    # update the x-axis label
#  facet_wrap(~gender)

# or

qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +    # update the x-axis label
  facet_wrap(~gender)

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)   # returns the number of female and male
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:more robust statistics

***

### Tenure
Notes:

```{r Tenure}
qplot(x = tenure, data = pf, color = I('black'), fill= I('#099DD9') )
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x = tenure/365, data = pf, 
      color = I('black'), fill= I('#099DD9') , 
      binwidth = 0.25 ) +
  scale_x_continuous(limits = c(0, 7),
                     breaks = seq(0, 7, 1))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(x = tenure/365, data = pf, 
      xlab = 'Number of years using Facbook',
      ylab = 'Number of users in sample',
      color = I('black'), fill= I('#099DD9') , 
      binwidth = 0.25 ) +
  scale_x_continuous(limits = c(0, 7),
                     breaks = seq(0, 7, 1))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(x = age, data = pf, 
      xlab = 'Age of Users',
      ylab = 'Number of users in sample',
      color = I('black'), fill= I('#099DD9') , 
      binwidth = 1 ) +
  scale_x_continuous(limits = c(10, 115),
                     breaks = seq(10, 115, 10)) 
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

```{r}
summary( pf$friend_count   )
summary( log10( pf$friend_count + 1)   )  #  + 1 to avoid to deal with - infinity
summary( sqrt( pf$friend_count + 1)   )
```


***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
#install.packages('gridExtra') 
library(gridExtra) 

# define individual plots
p1 <- qplot(x = friend_count, data = pf)
p2 <- qplot(x = friend_count, data = pf) +          # log10(friend_count+1)
  scale_x_log10()
p3 <- qplot(x = friend_count, data = pf) +          # sqrt(friend_count)
  scale_x_sqrt()
# arrange plots in grid
grid.arrange(p1, p2, p3, ncol=1)

# alternative:
# p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
# p2 <- p1 + scale_x_log10()
# p3 <- p1 + scale_x_sqrt()
#grid.arrange(p1, p2, p3, ncol=1)
```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)), 
      xlab = 'Friend count',
      ylab = 'Proportion of users with that friend count',
      color = gender , 
      binwidth = 10,
      geom = 'freqpoly') +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) 
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
qplot(x = www_likes, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)), 
      xlab = 'Likes on the WWW',
      ylab = 'Proportion of users with that number of likes',
      color = gender , 
      binwidth = 0.25,
      geom = 'freqpoly') +
  scale_x_continuous(limits = c(0, 10),
                     breaks = seq(0, 10, 1)) + 
  scale_x_log10()

# number of total likes per gender
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(x =gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot')
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x =gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot', ylim = c(0,1000))
# OR
qplot(x =gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') +
  scale_y_continuous(limits = c(0,1000))
# OR MORE ACCURATELY
qplot(x =gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0,1000))    # used so the summary output matches the box plot

by(pf$friend_count, pf$gender, summary)
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(x =gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0,250))
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
qplot(x =gender, y = friendships_initiated, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0,150))

by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)
qplot(data = pf, x = mobile_likes)+
    scale_x_continuous(limits = c(0, 10),
                     breaks = seq(0, 10, 1))

summary(pf$mobile_likes > 0)

mobile_check_in <- NA     # creating a new variable
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

# percent of check in using mobile: 65%
sum(pf$mobile_check_in == 1) / length(pf$mobile_check_in)
```

Response:

***

### Analyzing One Variable
Reflection:
look at the data distribution to find insightful information
look for outliers and missing values
plotting the data helps a lot, conveys more information
use data transformation can be very helpful too

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!
