Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots

Load data:

```{r}
getwd()
setwd('C:/Users/Nicolas/Desktop/Projets Tech/Data Analysis with R/')
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
names(pf)
```


Notes:

```{r Scatterplots}
library(ggplot2)
qplot(x= age, y = friend_count ,data = pf)

```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

ggplot(aes(x= age, y = friend_count) , data = pf) + 
  geom_point( alpha=1/20)    # set the transparency of the points using the alpha parameter (to address overplotting)
                             # use geom_point to create a scatterplot

```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x= age, y = friend_count) , data = pf) + 
  geom_jitter(alpha=1/20) +   # add some noise to have a more intuitive representation
  xlim(13,90)                # between age 13 to 90, to remove some anomalies
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x= age, y = friend_count) , data = pf) + 
  geom_point(alpha=1/20) +   # add some noise to have a more intuitive representation
  xlim(13,90) +               # between age 13 to 90, to remove some anomalies
  coord_trans(y = 'sqrt')
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x= age, y = friend_count) , data = pf) + 
  geom_point(alpha=1/20) +   # back to geom_point
  xlim(13,90) +               # between age 13 to 90, to remove some anomalies
  coord_trans(y = 'sqrt')
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x= age, y = friend_count) , data = pf) + 
  geom_point(alpha=1/20, position = position_jitter(h=0)) +   # add some noise, but need to be careful with negative friend count
  xlim(13,90) +               # between age 13 to 90, to remove some anomalies
  coord_trans(y = 'sqrt')
```

***

### Relationship between friends initiated  vs age
Notes:

```{r Alpha and Jitter}
ggplot(aes(x= age, y = friendships_initiated) , data = pf) + 
  geom_jitter(alpha=1/10, position = position_jitter(h=0)) +   # add some noise, but need to be careful with negative number
  xlim(13,90) +               # between age 13 to 90, to remove some anomalies
  coord_trans(y = 'sqrt')
```


***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
#install.packages('dplyr')
library(dplyr)                  # allows to split up a dataframe and apply functions

#most important functions are:
#filter()
#group_by()
#mutate()
#arrange()

age_groups <- group_by(pf, age)     # first groups data by age

pf.fc_by_age <- summarise(age_groups,                 # then summarizes
          friend_count_mean = mean(friend_count),     # and creates 3 new variables
          friend_count_median = median(friend_count),
          n = n()                                      # number of items in each group
          )

pf.fc_by_age <- arrange(pf.fc_by_age, age)                 # and sort by age

pf.fc_by_age
```

### Conditional Means
Notes:

```{r Conditional Means}
#install.packages('dplyr')
library(dplyr)                  # allows to split up a dataframe and apply functions

# alternatively, it is possible to chain the commands:
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),     # and creates 3 new variables
          friend_count_median = median(friend_count),
          n = n() ) %>%
  arrange(age) 

pf.fc_by_age
```



Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x= age, y = friend_count_mean) , data = pf.fc_by_age) + 
  geom_line()              # creates a line graph
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x= age, y = friend_count) , data = pf) + 
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000)) +  # used to zoom in (instead of xlim layer)
  #coord_trans(y = 'sqrt') +    either coord_trans or coord_cartesian
  geom_point(alpha=1/20, 
             position = position_jitter(h=0),
             color = 'orange') +   # add some noise, but need to be careful with negative friend count
  geom_line(stat = 'summary', fun.y = mean) +    # overlay the mean (= 'conditional mean')
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.1),
            linetype = 2, color ='blue') + #overlay the 10th percentile (the first 10%)
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.5),
            color ='blue') + #overlay the median (the first 50%)
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.9),
            linetype = 2, color ='blue')   #overlay the 90th percentile (the first 90%)
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor(pf$age, pf$friend_count)
#or
cor.test(pf$age, pf$friend_count, method="pearson")
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with( pf[ pf$age <= 70 ,]  , cor.test(age, friend_count, method="pearson"))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes: scatterplot of likes_received (y) vs. www_likes_received (x)

```{r}
ggplot(aes(x= www_likes_received, y = likes_received) , data = pf) + 
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, probs = 0.95)) +   # 95% percentile on x variable
  ylim(0, quantile(pf$likes_received, probs = 0.95))       # 95% percentile on y variable
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x= www_likes_received, y = likes_received) , data = pf) + 
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, probs = 0.95)) +   # 95% percentile on x variable
  ylim(0, quantile(pf$likes_received, probs = 0.95)) +       # 95% percentile on y variable
  geom_smooth(method = 'lm', color = 'red')      # adding linear regression
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
#cor(pf$www_likes_received,pf$likes_received)     # provides a number
cor.test(pf$www_likes_received,pf$likes_received)  # provides a report
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('car',dependencies=TRUE)
#install.packages('alr3',dependencies=TRUE)
#install.packages('nlme')
library(car)
library(alr3)
data(Mitchell)
?Mitchell
```

Create your plot!

```{r Temp vs Month}
library(ggplot2)
ggplot(aes(x= Mitchell$Month, y= Mitchell$Temp) , data = Mitchell) + 
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0

b. What is the actual correlation of the two variables?
(Round to the thousandths place)
0.05

```{r Noisy Scatterplots}
cor.test(Mitchell$Month,Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x= Mitchell$Month, y= Mitchell$Temp) , data = Mitchell) + 
  geom_point() +
  scale_x_discrete(breaks = seq(0,203,11))   # making month discrete
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
names(pf)

ggplot(aes(x= age, y= friend_count_mean) , data = pf.fc_by_age) + 
  geom_line() 

head(pf.fc_by_age, 10)
pf.fc_by_age[17:19,]
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- pf$age + (12 - pf$dob_month) / 12
```

Programming Assignment
```{r Programming Assignment}
library(dplyr)                  # allows to split up a dataframe and apply functions

age_with_months_groups <- group_by(pf, age_with_months)     # first groups data by age_with_months

pf.fc_by_age_months <- summarise(age_with_months_groups,    # then summarizes
          friend_count_mean = mean(friend_count),     # and creates 3 new variables
          friend_count_median = median(friend_count),
          n = n()                                      # number of items in each group
          )

pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)  # and sort by age w/ months

pf.fc_by_age_months
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x= age_with_months, y = friend_count_mean) , data = pf.fc_by_age_months) + 
  geom_line()  +
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 450))
```

***

### Comparing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x= age, y = friend_count_mean) , 
             data = subset(pf.fc_by_age, age<71)) + 
  geom_line()

p2 <- ggplot(aes(x= age_with_months, y = friend_count_mean) , 
             data = subset(pf.fc_by_age_months, age_with_months<71)) + 
  geom_line() 

p3 <- ggplot(aes(x= round(age/5)*5, y = friend_count) , 
             data = subset(pf, age<71)) + 
  geom_line(stat = 'summary', fun.y = mean) 

library(gridExtra)
grid.arrange(p2,p1,p3,ncol=1)
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x= age, y = friend_count_mean) , 
             data = subset(pf.fc_by_age, age<71)) + 
  geom_line() +
  geom_smooth()              # adding a Local Regression plot (LOESS)

p2 <- ggplot(aes(x= age_with_months, y = friend_count_mean) , 
             data = subset(pf.fc_by_age_months, age_with_months<71)) + 
  geom_line() +
  geom_smooth()             # adding a Local Regression plot (LOESS)

p3 <- ggplot(aes(x= round(age/5)*5, y = friend_count) , 
             data = subset(pf, age<71)) + 
  geom_line(stat = 'summary', fun.y = mean) 

library(gridExtra)
grid.arrange(p2,p1,p3,ncol=1)
```


***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

